---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Estimating observer error and steelhead redd abundance using a modified Gaussian area-under-the-curve framework"
Author <- "Murdoch, A. R. and Herring, C. J. and Frady, C. H. and See, K. E. and Jordan, C. E."
Year <- 2018
Journal <- "Canadian Journal of Fisheries and Aquatic Sciences"
```

# SthdReddObsError

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio)

This repository contains the data and code for our paper:

> `r Author`, (`r Year`). _`r Title`_. `r Journal`  <https://doi.org/10.1139/cjfas-2017-0335>

It also contains data and code demonstrating how this model has been used to estimate redds, using predictions of redd observer error. These estimates of redds are translated to estimates of spawners, and divided into natural origin and hatchery origin spawners, based on data from a separate [project](https://github.com/KevinSee/DabomPriestRapidsSthd) using PIT tagged steelhead to estimates escapement to various areas in the Upper Columbia. The reports are contained in the "outoing/reports" folder that is available after cloning or downloading this compendium. 

### How to cite

Please cite this compendium as:

> See, K.E., (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.1139/cjfas-2017-0335>

### How to download or install

You can download the compendium as a zip from from this URL: <https://github.com/KevinSee/SthdReddObsError/archive/master.zip>

Or you can install this compendium as an R package, SthdReddObsError, from GitHub with:


```{r gh-installation, eval = FALSE}
# install.packages("devtools")
remotes::install_github("KevinSee/SthdReddObsError")
```


### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

