---
title: "Estimates of Wenatchee Steelhead Redds in 2019"
author:
  - Kevin See:
      email: Kevin.See@biomark.com
      institute: [biomark]
      correspondence: true
institute:
  - biomark: Biomark, Inc.
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 6
      fig_width: 6
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
        smooth_scroll: yes
      theme: simplex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 5
      fig_width: 6
      toc: yes
      includes:
        in_header: ../templates/header_ABS.tex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      fig_height: 4
      fig_width: 6
      toc: yes
      reference_docx: "../templates/ReportTemplate.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- packages.bib
- references.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  This report contains estimates of total steelhead redds in the Wenatchee, after accounting for observer bias.
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights.
---


<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

```{r packages}
library(tidyverse)
library(lubridate)
library(janitor)
library(magrittr)
library(msm)
library(knitr)
library(kableExtra)
library(english)
library(SthdReddObsError)

theme_set(theme_bw())

options(knitr.kable.NA = '-')
# options(knitr.table.format = "pandoc")

# setwd('analysis/paper')
```

```{r package-bibtex, eval = F}
knitr::write_bib(c("tidyverse", 
                   "lubridate", 
                   "janitor", 
                   "knitr",
                   "magrittr",
                   "msm"),
                 file = 'packages.bib')
```


```{r load-data}
redd_df = read_rds('../data/derived_data/wen_2019.rds')
org_data = read_rds('../data/derived_data/original_data.rds')
net_err_mod = read_rds('../data/derived_data/net_error_model.rds')

# what year is this report for?
yr = unique(year(redd_df$SurveyDate))
```

```{r calc-total-redds}
#-------------------------------------
# set some thresholds
# minimum number of total redds observed
min_redds = 2
# minimum number of weeks with at least one new redd observed
min_non0_wks = 3
#-------------------------------------

redd_results = redd_df %>%
  group_by(River, Reach, Index, SurveyType) %>%
  summarise(n_weeks = n(),
            n_non0_wks = sum(NewRedds > 0, na.rm=T),
            tot_feat = sum(NewRedds, na.rm=T),
            err_est = mean(NetError), 
            err_se = mean(NetErrorSE)) %>%
  ungroup() %>%
  mutate(err_se = ifelse(is.na(err_est), 0, err_se),
         err_est = ifelse(is.na(err_est), 1, err_est)) %>%
  full_join(redd_df %>%
              group_by(River, Reach, Index, SurveyType) %>%
              nest()) %>%
  mutate(gauc_list = map(data,
                         .f = function(x) {
                           mod_df = x %>%
                             select(redds = NewRedds) %>%
                             mutate(day = 1:n())
                           
                           res_list = fit_gauc(data = mod_df,
                                               v = if_else(is.na(mean(x$NetError, na.rm = T)),
                                                           1,
                                                           mean(x$NetError, na.rm = T)),
                                               v_se = if_else(is.na(mean(x$NetError, na.rm = T)),
                                                              0,
                                                              mean(x$NetErrorSE, na.rm = T)))
                           return(res_list)
                         })) %>%
  mutate(converged = map_lgl(gauc_list,
                             .f = function(x) {
                               x$model$converged
                             })) %>%
  mutate(redd_est = map_dbl(gauc_list,
                            .f = 'E'),
         redd_se = map_dbl(gauc_list,
                           .f = 'E_se')) %>%
  mutate(GAUC = if_else(!converged | n_non0_wks < min_non0_wks | tot_feat < min_redds,
                        F, T)) %>%
  rowwise() %>%
  mutate(redd_est = if_else(!GAUC,
                            tot_feat / err_est,
                            redd_est),
         redd_se = if_else(!GAUC,
                           msm::deltamethod(~ x1 / x2,
                                            mean = c(tot_feat, err_est),
                                            cov = diag(c(0, err_se)^2)),
                           redd_se)) %>%
  ungroup() %>%
  mutate_at(vars(redd_est),
            list(round_half_up))

```


\newpage

# Introduction

Redd counts are an established method to provide an index of adult spawners [@Gallagher2007]. In the Wenatchee and Methow subbasins, index reaches are surveyed weekly during the steelhead spawning season (`r paste(format(range(redd_df$SurveyDate), '%b %d, %Y'), collapse = ' - ')`) and non-index reaches are surveyed once during the peak spawning period. The goal of this work is to:

1. Predict observer net error, using the model described in @Murdoch2018.
1. Use estimates of observer net error rates and the mean survey interval to estimate the number of redds in each index reach, using a Gaussian area under the curve (GAUC) technique described in @Millar2012.
1. Estimate the total number of redds in the non-index reaches by adjusting the observed counts with the estimated net error.
1. Sum the total number of estimated redds for the entire Wenatchee subbasin.

# Methods

## Net Error Model

The model for observer net error (observed redd counts / true number of redds) is fully described in @Murdoch2018. It uses covariates of the log of observer experience, mean discharge, the observed redd density and mean thalweg CV as a proxy for channel complexity. After normalizing these covariates, the model coefficients are shown in Table \@ref(tab:model-summ). 

```{r model-summ}
summary(net_err_mod)$coefficients %>%
  as_tibble(rownames = 'Covariate') %>%
  mutate(Covariate = recode(Covariate, 
                           'MeanDischarge' = 'Mean Discharge',
                           'MeanThalwegCV' = 'Mean Thalweg CV',
                           'NaiveDensity_km' = 'Obs. Redd Density',
                           'ExpSpTotal_log' = 'Log Surveyor Exp.')) %>%
  arrange(desc(abs(Estimate))) %>%
  select(1:3) %>%
  kable(digits = 3,
        booktabs = T,
        caption = "Net error model covariates and coefficients.")
```

## Data

Redd counts were conducted on a nearly weekly basis for index reaches, and once during the peak spawning period for non-index reaches. The covariates in the observer error model of @Murdoch2018 were collected during each survey. They were compared with the covariates contained in the model data set, as well as the estimates of net error (Figure \@ref(fig:covariate-comparison)). 

```{r covariate-comparison, fig.cap = "Net error covariate values from the original study the predicted reaches in this report."}
redd_df %>%
  select(NetError,
         one_of(attr(net_err_mod$terms, 'term.labels'))) %>%
  mutate(Source = paste(unique(year(redd_df$SurveyDate)), 'Data')) %>%
  bind_rows(org_data %>%
              select(NetError,
                     one_of(attr(net_err_mod$terms, 'term.labels'))) %>%
              mutate(Source = 'Model Data')) %>%
  gather(variable, value, -Source) %>%
  mutate(variable = recode(variable, 
                           'NetError' = 'Net Error',
                           'MeanDischarge' = 'Mean Discharge',
                           'MeanThalwegCV' = 'Mean Thalweg CV',
                           'NaiveDensity_km' = 'Obs Redd Density',
                           'ExpSpTotal_log' = 'Log Surveyor Exp.')) %>%
  mutate(variable = factor(variable,
                           levels = c('Mean Thalweg CV',
                                      'Obs Redd Density',
                                      'Log Surveyor Exp.',
                                      'Mean Discharge',
                                      'Net Error'))) %>%
  ggplot(aes(x = Source,
             y = value,
             fill = Source)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = 'bottom') +
  facet_wrap(~ variable,
             scales = 'free_y')

```

Those covariates in the observer error model were collected during each survey in `r yr`, and predictions of net error were made for each survey. From these survey specific estimates of net error, a mean and standard error of net error was calculated for each reach. The standard deviation was calculated by taking the square root of the sum of the squared standard errors for all predictions within a reach.

## Estimating Redds

Estimates of total redds were made for each index reach with a minimum of `r as.character(as.english(min_redds))` redds and at least `r as.character(as.english(min_non0_wks))` weeks with at least one new redd found using the GAUC model described in @Millar2012 and @Murdoch2018. The GAUC model was developed with spawner counts in mind. As it is usually infeasible to mark every individual spawner, only total spawner counts can be used, and an estimate of average stream life must be utilized to translate total spawner days to total unique spawners. However, in adapting this for redd surveys, we note that individual redds can be marked, and therefore the GAUC model can be fit to new redds only. The equivalent of stream life thus the difference between survey numbers which can be fixed at 1.

For non-index reaches, which were surveyed only once during peak spawning, the estimate of total redds was calculated by dividing the observed redds by the estimate of net error associated with that survey. This assumes that no redds were washed out before the non-index survey, and that no new redds appeared after that survey. As the number of redds observed in the non-index reaches ranged from `r paste(range(redd_df$NewRedds[redd_df$Index == 'N']), collapse = ' to ')`, any violoation of this assumption should not affect the overall estimates very much. Any index reaches that did not meet the thresholds described above were treated as non-index reaches, and the total observed redds in those reaches were divided by the mean estimate of net error for each reach.

\newpage

# Results

## Redd estimates

The estimated net error, observed redds and estimates of redds are shown in Table \@ref(tab:redd-est-rch). Plots of the new redds and the GAUC fit to those data are shown in Figure \@ref(fig:gauc-plots).

```{r redd-est-rch}
redd_results %>%
  mutate(Type = recode(Index,
                       "Y" = 'Index',
                       "N" = "Non-Index")) %>%
  mutate(err_cv = err_se / err_est,
         redd_cv = redd_se / redd_est) %>%
  bind_rows(redd_results %>%
              summarise(River = "Total",
                        tot_feat = sum(tot_feat, na.rm = T),
                        redd_est = sum(redd_est, na.rm = T),
                        redd_se = sqrt(sum(redd_se^2, na.rm = T)),
                        redd_cv = redd_se / redd_est)) %>%
  select(River, Reach, Type, 
         # GAUC,
         `Net Error` = err_est,
         `Net Error CV` = err_cv,
         `Observed Redds ` = tot_feat,
         `Estimated Redds` = redd_est,
         `Redds CV` = redd_cv) %>%
  kable(digits = c(rep(NA, 3), rep(3, 2), 1, 1, 3),
        format.args = list(drop0trailing = T),
        booktabs = T,
        caption = "Estimates of mean net error and redds for each reach.") %>%
  kable_styling(latex_options = 'striped') %>%
  row_spec(row = nrow(redd_results),
           hline_after = T) %>%
  row_spec(row = nrow(redd_results) + 1, 
           bold = T)
  
```

```{r gauc-plots, fig.cap = "Plots of observed redd counts (black dots) through time for each qualifying index reach, and the fitted curve from the GAUC model (blue line) with associated uncertainty (gray)."}
redd_results %>%
  filter(GAUC) %>%
  select(River:SurveyType, data) %>%
  unnest(cols = c(data)) %>%
  group_by(River, Reach, Index) %>%
  mutate(survey = 1:n()) %>%
  ungroup() %>%
  ggplot(aes(x = survey, 
             y = NewRedds,
             group = Reach)) +
  geom_point() +
  facet_wrap(~ Reach, 
             scales='free',
             ncol = 2) +
  stat_smooth(method = 'glm', 
              formula = y ~ x + I(x^2), 
              method.args = list(family = quasipoisson), 
              fullrange = T, 
              se = T) + 
  labs(x = 'Survey Number', 
       y = 'New Redds', 
       title='GAUC Reaches') +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  theme(strip.text = element_text(size=12))

```

# Discussion

Most of the covariates collected in `r yr` were within the range of those in the model data set from @Murdoch2018, leading to estimates of net error in a very similar range to the model dataset (Figure \@ref(fig:covariate-comparison)). However, most reaches did not meet the minimum thresholds of number of observed redds or number of weeks with at least one new redd observed, so we used the GAUC method in only `r as.character(as.english(sum(redd_results$GAUC)))` reaches. 

# Acknowledgements

The data for this report was collected by Washington Department of Fish and Wildlife. Development of the observer error model was done in collaboration with Andrew Murdoch, WDFW.

<!-- The following line inserts a page break  -->
\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

<!-- \newpage -->

<!-- ### Colophon -->

<!-- This report was generated on `r Sys.time()` using the following computational environment and dependencies:  -->

<!-- ```{r colophon, cache = FALSE} -->
<!-- # which R packages and versions? -->
<!-- if ("devtools" %in% installed.packages()) devtools::session_info() -->
<!-- ``` -->

<!-- The current Git commit details are: -->

<!-- ```{r} -->
<!-- # what commit is this file at?  -->
<!-- if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())   -->
<!-- ``` -->
