---
title: "Estimates of Methow Steelhead Redds and Spawners"
subtitle: "Spawn Year 2021"
author:
  - Kevin See:
      email: Kevin.See@dfw.wa.gov
      institute: [wdfw]
      correspondence: true
institute:
  - wdfw: Washington Department of Fish & Wildlife
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    wdfwTemplates::wdfw_html_format2:
      fig_caption: yes
      fig_height: 6
      fig_width: 6
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
        smooth_scroll: yes
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 6
      fig_width: 6
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
        smooth_scroll: yes
      theme: simplex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 5
      fig_width: 6
      toc: yes
      includes:
        in_header: ../templates/header_wdfw.tex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      fig_height: 4
      fig_width: 6
      toc: yes
      reference_docx: "../templates/ReportTemplate.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- packages.bib
- references.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  This report contains estimates of total steelhead redds in the Methow, after accounting for observer bias. It also includes estimates of spawners, as well as prespawn mortality.
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights.
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

```{r packages}

# load these packages
library(tidyverse)
library(lubridate)
library(janitor)
library(magrittr)
library(msm)
library(knitr)
library(kableExtra)
library(english)
library(UCSthdReddObsErr)
library(here)

theme_set(theme_bw())

# knitr options
options(knitr.kable.NA = '-')

# when knitting to Word, use this
# what kind of document is being created?
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

if(doc.type == 'docx') {
  options(knitr.table.format = "pandoc")
}

```

```{r package-bibtex, eval = F}
knitr::write_bib(c("tidyverse", 
                   "lubridate", 
                   "janitor", 
                   "knitr",
                   "magrittr",
                   "msm"),
                 file = 'packages.bib')
```


```{r load-data}
# what year is this report for?
yr = 2021

# load data for this year
load(here('analysis/data/derived_data',
          paste0('met_', yr, '.rda')))

# load data from model fitting
org_data = read_rds(here('analysis/data/derived_data/two_obs_original_data.rds'))
net_err_mod = read_rds(here('analysis/data/derived_data/two_obs_net_error_model.rds'))

# check we have the right year
if(unique(year(redd_df$SurveyDate)) != yr) {
  print("Incorrect year")
}

# how many fish removed this year?
data("removal_df")
rem_df = removal_df %>%
  filter(Subbasin == "Methow",
         Year == yr) %>%
  select(-Year,
         -Subbasin)


```

```{r calc-total-redds}
#-------------------------------------
# set some thresholds
# minimum number of total redds observed
min_redds = 2
# minimum number of weeks with at least one new redd observed
min_non0_wks = 3
#-------------------------------------

redd_results = redd_df %>%
  group_by(River, Reach, Index, SurveyType) %>%
  summarise(n_weeks = n(),
            n_non0_wks = sum(NewRedds > 0, na.rm=T),
            tot_feat = sum(NewRedds, na.rm=T),
            err_est = mean(NetError[VisibleRedds > 0]), 
            err_se = mean(NetErrorSE[VisibleRedds > 0])) %>%
  ungroup() %>%
  mutate(err_se = ifelse(is.na(err_est), 0, err_se),
         err_est = ifelse(is.na(err_est), 1, err_est)) %>%
  full_join(redd_df %>%
              group_by(River, Reach, Index, SurveyType) %>%
              nest()) %>%
  mutate(gauc_list = map(data,
                         .f = function(x) {
                           mod_df = x %>%
                             select(redds = NewRedds) %>%
                             mutate(day = 1:n())

                           v_vec = x %>%
                             filter(VisibleRedds > 0) %>%
                             summarise_at(vars(NetError, NetErrorSE),
                                          list(mean),
                                          na.rm = T) %>%
                             mutate(NetErrorSE = if_else(is.na(NetError),
                                                         0,
                                                         NetErrorSE),
                                    NetError = if_else(is.na(NetError),
                                                       1,
                                                       NetError))
                           
                           # # if you want to assume no observer error
                           # v_vec = tibble(NetError = 1,
                           #                NetErrorSE = 0)
                           
                           res_list = fit_gauc(data = mod_df,
                                               v = v_vec$NetError,
                                               v_se = v_vec$NetErrorSE)
                           return(res_list)
                         })) %>%
  mutate(converged = map_lgl(gauc_list,
                             .f = function(x) {
                               x$model$converged
                             })) %>%
  mutate(redd_est = map_dbl(gauc_list,
                            .f = 'E'),
         redd_se = map_dbl(gauc_list,
                           .f = 'E_se')) %>%
  mutate(GAUC = if_else(!converged | n_non0_wks < min_non0_wks | tot_feat < min_redds,
                        F, T)) %>%
  rowwise() %>%
  mutate(redd_est = if_else(!GAUC,
                            tot_feat / err_est,
                            redd_est),
         redd_se = if_else(!GAUC,
                           msm::deltamethod(~ x1 / x2,
                                            mean = c(tot_feat, err_est),
                                            cov = diag(c(0, err_se)^2)),
                           redd_se)) %>%
  ungroup() %>%
  mutate_at(vars(redd_est),
            list(round_half_up))

```

```{r tribs-below-arrays, eval = F}
redd_results %<>%
  bind_rows(redds_below_arrays %>%
              mutate(Index = "N",
                     err_est = 1,
                     err_se = 0,
                     tot_feat = redd_est,
                     GAUC = F))
```


```{r calc-strm-totals}
mainstm_est_naive = redd_results %>%
  filter(str_detect(Reach,
                    "MRW[[:digit:]]",
                    negate = F)) %>%
  mutate(Location = "Lower Methow") %>%
  select(River, Index, Location, Reach, tot_feat, redd_est, redd_se) %>%
  group_by(River, Location, Index) %>%
  nest() %>%
  rename(redd_tot = data) %>%
  mutate(n_rchs = map_int(redd_tot,
                          .f = function(x) {
                            n_distinct(x$Reach)
                          })) %>%
  # get correlations between reaches within a stream
  left_join(redd_df %>%
              mutate(Location = "Lower Methow") %>%
              group_by(River, Location, Index) %>%
              nest()) %>%
  # filter(n_rchs > 1) %>%
  mutate(cor_mat = map(data,
                       .f = function(x) {
                         cor_mat = try(x %>%
                                         group_by(Reach) %>%
                                         mutate(Survey = 1:n()) %>%
                                         ungroup() %>%
                                         select(Reach, Survey, NewRedds) %>%
                                         spread(Reach, NewRedds) %>%
                                         select(-Survey) %>%
                                         cor(use = "pairwise",
                                             # method = 'kendall')
                                             method = 'spearman'),
                                       silent = T)
                         
                         if(class(cor_mat) == 'try-error') {
                           return(as.numeric(1))
                         }else{
                           
                           cor_mat[is.na(cor_mat)] = 0
                           diag(cor_mat) = 1
                           return(cor_mat)
                         }
                       })) %>%
  # select(-data) %>%
  mutate(strm_obs = map_dbl(redd_tot,
                            .f = function(x) {
                              sum(x$tot_feat, na.rm = T)
                            }),
         strm_est = map_dbl(redd_tot,
                            .f = function(x) {
                              sum(x$redd_est, na.rm = T)
                            }),
         strm_se = map_dbl(redd_tot,
                            .f = function(x) {
                              sqrt(sum(x$redd_se, na.rm = T)^2)
                            })) %>%
  ungroup()

mainstm_est = mainstm_est_naive %>%
  filter(n_rchs > 1,
         Index == "Y") %>%
  # use correlations between reaches to get appropriate standard error
  mutate(strm_se = map2_dbl(redd_tot, 
                            cor_mat,
                            .f = function(x, y) {
                              
                              deltamethod(as.formula(paste('~', paste0('x', 1:nrow(x), collapse='+'))),
                                          mean = x$redd_est,
                                          cov = diag(x$redd_se) %*% y %*% diag(x$redd_se))
                            })) %>%
  bind_rows(anti_join(mainstm_est_naive,
                      .,
                      by = c('River', 'Index', 'Location', 'n_rchs'))) %>%
  arrange(Location, River, Index) %>%
  select(-redd_tot, -data, -cor_mat) %>%
  mutate(strm_cv = strm_se / strm_est)
```

```{r spawner-estimate}
low_met_spwn <- redd_results %>%
  filter(River == 'Methow') %>%
  mutate(Location = "Lower Methow") %>%
  select(River, Location, Reach, Index, redd_est, redd_se) %>%
  left_join(met_sex_tab %>%
              select(Location, starts_with('fpr'))) %>%
  left_join(met_prop_origin) %>%
  rowwise() %>%
  mutate(Tot_Spawners = redd_est * fpr,
         Org_Spawners = Tot_Spawners * prop,
         Org_SE = if_else(Tot_Spawners > 0,
                          deltamethod(~ x1 * x2 * x3,
                                      mean = c(redd_est,
                                               fpr,
                                               prop),
                                      cov = diag(c(redd_se,
                                                   fpr_se,
                                                   prop_se)^2)),
                          NA_real_)) %>%
  ungroup() %>%
  mutate(Reach = factor(Reach,
                        levels = c(paste0('MRW', 1:8)))) %>%
  select(Reach, Location, Type = Index, 
         Origin,
         starts_with("Org"))

trib_spwn <- escp_met %>%
  filter(!Area %in% c("Methow_all",
                      "Lower Methow",
                      "Methow Fish Hatchery",
                      "Spring Creek")) %>%
  rename(Group = Origin) %>%
  full_join(tibble(Group = "Hatchery",
                   Origin = c("HOR-SN",
                              "HOR-C",
                              NA_character_)) %>%
              bind_rows(tibble(Group = "Natural",
                               Origin = "W"))) %>%
  inner_join(met_prop_origin,
             by = c("Area" = "Location",
                    "Origin")) %>%
  mutate(Origin = factor(Origin,
                         levels = levels(met_prop_origin$Origin))) %>%
  group_by(Location = Area, Group) %>%
  mutate(tot_tags = sum(n_tags),
         prop = n_tags / tot_tags,
         prop_se = sqrt((prop * (1 - prop)) / tot_tags)) %>%
  rowwise() %>%
  mutate(Org_Spawners = estimate * prop,
         Org_SE = deltamethod(~ x1 * x2,
                              mean = c(estimate, prop),
                              cov = diag(c(se,
                                           prop_se)^2))) %>%
  ungroup() %>%
  mutate(Type = "Trib") %>%
  select(Location, Type, 
         Origin,
         starts_with("Org"))

spwn_rch <- low_met_spwn %>%
  bind_rows(trib_spwn) %>%
  mutate(Location = factor(Location,
                           levels = levels(met_prop_origin$Location))) %>%
  mutate(across(Location,
                fct_drop)) %>%
  arrange(Location, Reach,
          Type)

# spwn_rch %>%
#   mutate(prnt_val = paste0(round_half_up(Org_Spawners),
#                            " (",
#                            round(Org_SE, 2),
#                            ")")) %>%
#   select(-starts_with("Org")) %>%
#   pivot_wider(names_from = Origin,
#               values_from = prnt_val,
#               names_sort = T)
```


\newpage

# Introduction

Redd counts are an established method to provide an index of adult spawners [@Gallagher2007]. In the Methow subbasin, index reaches are surveyed weekly during the steelhead spawning season (`r paste(format(range(redd_df$SurveyDate), '%b %d, %Y'), collapse = ' - ')`) and non-index reaches are surveyed once during the peak spawning period. The goal of this work is to:

1. Predict observer net error, using the model described in @Murdoch2018.
1. Use estimates of observer net error rates and the mean survey interval to estimate the number of redds in each index reach, using a Gaussian area under the curve (GAUC) technique described in @Millar2012 and @Murdoch2018.
1. Estimate the total number of redds in the non-index reaches by adjusting the observed counts with the estimated net error where possible.
1. Sum the total number of estimated redds for the entire Methow subbasin.

# Methods

## Net Error Model

The net error ($NE$) for a reach $i$ is defined as

$$
NE_i = \frac{F_i}{V_i}
$$

where $F_i$ is the number of redds the surveyor reported and $V_i$ is the true number of redds in the reach. Therefore, if we have an estimate of net error ($\hat{NE_i}$), we can calculate the true number of redds based on that estimate and the number of redds the surveyor reported:

\begin{equation}
  V_i = \frac{F_i}{\hat{NE_i}}
  (\#eq:net-error)
\end{equation}

The model for observer net error is fully described in @Murdoch2018. It uses covariates of the log of observer experience, mean discharge, the observed redd density and mean thalweg CV as a proxy for channel complexity. After normalizing these covariates, the model coefficients are shown in Table \@ref(tab:model-summ). The response, net error, is scaled such that estimates of net error less than 1 suggest more errors of omission, while estimates greater than 1 suggest more errors of commission. An estimate of net error equal to one would indicate the observed count equals the true number of redds.

```{r model-summ}
summary(net_err_mod)$coefficients %>%
  as_tibble(rownames = 'Covariate') %>%
  mutate(Covariate = recode(Covariate, 
                           'MeanDischarge' = 'Mean Discharge',
                           'MeanThalwegCV' = 'Mean Thalweg CV',
                           'NaiveDensity_km' = 'Obs. Redd Density',
                           'ExpSpTotal_log' = 'Log Surveyor Exp.')) %>%
  arrange(desc(abs(Estimate))) %>%
  select(1:3) %>%
  kable(digits = 3,
        booktabs = T,
        caption = "Net error model covariates and coefficients.") %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped",
                                  "hold_position"))
```

## Data

Redd counts were conducted on a nearly weekly basis for index reaches, and once during the peak spawning period for non-index reaches. The covariates in the observer error model of @Murdoch2018 were collected during each survey. The covariate of mean thalweg CV was calculated based on all measurements taken within a reach across years (assuming this covariate does not vary through time within a reach). They were compared with the covariates contained in the model data set, as well as the estimates of net error (Figure \@ref(fig:covariate-comparison)). 

```{r covariate-comparison, fig.cap = "Net error covariate values from the original study the predicted reaches in this report."}
redd_df %>%
  select(NetError,
         one_of(attr(net_err_mod$terms, 'term.labels'))) %>%
  mutate(Source = paste(unique(year(redd_df$SurveyDate)), 'Data')) %>%
  bind_rows(org_data %>%
              select(NetError,
                     one_of(attr(net_err_mod$terms, 'term.labels'))) %>%
              mutate(Source = 'Model Data')) %>%
  gather(variable, value, -Source) %>%
  mutate(variable = recode(variable, 
                           'NetError' = 'Net Error',
                           'MeanDischarge' = 'Mean Discharge',
                           'MeanThalwegCV' = 'Mean Thalweg CV',
                           'NaiveDensity_km' = 'Obs Redd Density',
                           'ExpSpTotal_log' = 'Log Surveyor Exp.')) %>%
  mutate(variable = factor(variable,
                           levels = c('Mean Thalweg CV',
                                      'Obs Redd Density',
                                      'Log Surveyor Exp.',
                                      'Mean Discharge',
                                      'Net Error'))) %>%
  ggplot(aes(x = Source,
             y = value,
             fill = Source)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Set1") +
  theme(legend.position = 'bottom') +
  facet_wrap(~ variable,
             scales = 'free_y')

```

Those covariates in the observer error model were collected during each survey in `r yr`, and predictions of net error were made for each survey. From these survey specific estimates of net error, a mean and standard error of net error was calculated for each reach. The standard deviation was calculated by taking the square root of the sum of the squared standard errors for all predictions within a reach.

## Estimating Redds

Use of the GAUC methodology was limited to index reaches with a minimum of `r as.character(as.english(min_redds))` redds and at least `r as.character(as.english(min_non0_wks))` weeks with at least one new redd found. For those reaches, we used the method described in @Millar2012 and @Murdoch2018. The GAUC model was developed with spawner counts in mind. As it is usually infeasible to mark every individual spawner, only total spawner counts can be used, and an estimate of average stream life must be utilized to translate total spawner days to total unique spawners. However, in adapting this for redd surveys, we note that individual redds can be marked, and therefore the GAUC model can be fit to new redds only. The equivalent of stream life thus the difference between survey numbers which can be fixed at 1. We applied the average net error from all the surveys when redds were visible, so as to not bias the estimate from early weeks when no redds were found, and the observed redd density was zero.

For non-index reaches, which were surveyed only once during peak spawning, the estimate of total redds was calculated by dividing the observed redds by the estimate of net error associated with that survey (Eq. \@ref(eq:net-error)). This assumes that no redds were washed out before the non-index survey, and that no new redds appeared after that survey. As the number of redds observed in the non-index reaches ranged from `r paste(range(redd_df$NewRedds[redd_df$Index == 'N']), collapse = ' to ')`, any violation of this assumption should not affect the overall estimates very much. Any index reaches that did not meet the thresholds described above were treated as non-index reaches, and the total observed redds in those reaches were divided by the mean estimate of net error for each reach.

When summing reach-scale estimates to obtain estimates at the stream scale, an attempt was made to incorporate the fact that the reaches within a stream are not independent. Estimates of correlation between the reaches within a stream were made based on weekly observed redds. This method may not be perfect, since spawners could use certain reaches preferentially at different times in the season, but it may be the best we can do. Because correlations are often quite high between reaches, this is a better alternative than to naively assume the standard errors between reaches are independent of one another. These correlation estimates were combined with estimates of standard error at the reach scale to calculate a covariance matrix for the reaches within each stream, which was used when summing estimates of total redds to estimate the standard error at the stream scale. Failure to incorporate the correlations between reaches would result in an underestimate of standard error at the stream scales. Different streams (and therefore reaches in different streams) were assumed to be independent.

## Estimating Spawners

Estimates of escapement to various tributaries in the Methow were made using a branching patch-occupancy model [@Waterhouse2020] based on PIT tag observations of fish tagged at Priest Rapids dam. All fish that escaped to the various tributaries were assumed to be spawners (i.e. prespawn mortality only occurs in the mainstem). 
<!-- Reaches below the PIT tag arrays in some tributaries were surveyed for redds, but we assumed there was no observer error in those reaches.  -->

To convert estimates of redds in mainstem areas into estimates of natural and hatchery spawners, the estimates of redds were multiplied by a fish per redd (FpR) estimate and then by the proportion of hatchery or wild fish. The fish per redd estimate was based on PIT tags from the branching patch-occupancy model observed to move into the lower mainstem Methow, but not into tributaries (between PTAGIS sites LMR and MRW). FpR was calculated as the ratio of male to female fish, plus 1. Reaches MRW1 - MRW8 are all in the lower mainstem Methow below Winthrop. Similarly, the proportion of hatchery and natural origin fish was calculated from the same group of PIT tags for areas in the mainstem, and the various tributaries. 
<!-- For reaches in tributaries below the tributary PIT tag array, FpR and proportion hatchery were calculated based on observed PIT tags moving into the lower regions of each tributary. -->

## Prespawn Mortality

After translating estimates of redds to estimates of spawners by origin, we can then compare the spawner estimates to escapement estimates made using PIT tags, and estimate a prespawn mortality rate. Taking the total PIT-tag based escapement estimate, by origin, to the Methow, we then subtract any fish removed for broodstock or surplus, as well as any deaths due to harvest (Table \@ref(tab:removals)), and then subtract the total estimate of spawners, including the tributaries, to provide an estimate of how many fish succumbed to prespawn mortality. Dividing that number by the total escapement estimate provides an estimate of the prespawn mortality rate, by origin across the entire Methow population. 

```{r removals}
rem_df %>%
  rename(Removed = rem) %>%
  select(-Area) %>%
  spread(Origin, Removed) %>%
  kable(caption = "Known number of fish removed at various locations, by origin.",
        booktabs = T) %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped"))
```


We can also compare estimates of escapement to the lower mainstem Methow (after subtracting the the fish removed there) to the total estimate of spawners in the same area This allows us to estimate prespawn mortality in the lower mainstem Methow, by origin. 

If any group had a higher estimates of spawners compared to escapement, we fixed our prespawn mortality estimate at 0, reflecting a very low level of prespawn mortality. There is uncertainty in both the escapement and spawner estimates, which could explain why these types of scenarios could arise. 

```{r}
ps_mort_df = escp_met %>%
  filter(Area == "Lower Methow") %>%
  left_join(rem_df %>%
              # filter(Area != 'Harvest') %>%
              select(-Source)) %>%
  mutate(across(rem,
                replace_na,
                0)) %>%
  mutate(escp = estimate - rem) %>%
  left_join(low_met_spwn %>%
              mutate(Origin = fct_recode(Origin,
                                         "Hatchery" = "HOR-SN",
                                         "Hatchery" = "HOR-C",
                                         "Natural" = "W")) %>%
              group_by(Location, Origin) %>%
              summarize(spwn_est = sum(Org_Spawners),
                        spwn_se = sqrt(sum(Org_SE^2)),
                        .groups = "drop"),
            by = c("Area" = "Location",
                   "Origin")) %>%
  mutate(ps_mort = (escp - spwn_est) / escp) %>%
  rowwise() %>%
  mutate(ps_mort_se = deltamethod(~ (x1 - x2) / x1,
                                  mean = c(escp, spwn_est),
                                  cov = diag(c(se, spwn_se)^2))) %>%
  ungroup()

```

```{r}
ps_mort_met = escp_met %>%
  filter(Area == 'Methow_all') %>%
  left_join(rem_df %>%
              group_by(Origin) %>%
              summarize(across(rem,
                               sum),
                        .groups = "drop")) %>%
  mutate(across(rem,
                replace_na,
                0)) %>%
  # mutate(rem = 0) %>%
  mutate(escp = estimate - rem) %>%
  left_join(low_met_spwn %>%
              mutate(Origin = fct_recode(Origin,
                                         "Hatchery" = "HOR-SN",
                                         "Hatchery" = "HOR-C",
                                         "Natural" = "W")) %>%
              group_by(Location, Origin) %>%
              summarize(Spawners = sum(Org_Spawners),
                        Spawners_SE = sqrt(sum(Org_SE^2)),
                        .groups = "drop") %>%
              bind_rows(escp_met %>%
                          filter(!Area %in% c("Methow_all",
                                              "Lower Methow",
                                              "Methow Fish Hatchery",
                                              "Spring Creek")) %>%
                          rename(Spawners = estimate,
                                 Spawners_SE = se)) %>%
              group_by(Origin) %>%
              summarize(spwn_est = sum(Spawners),
                        spwn_se = sqrt(sum(Spawners_SE^2)),
                        .groups = "drop")) %>%
  mutate(ps_mort = (escp - spwn_est) / escp) %>%
  rowwise() %>%
  mutate(ps_mort_se = deltamethod(~ (x1 - x2) / x1,
                                  mean = c(escp, spwn_est),
                                  cov = diag(c(se, spwn_se)^2))) %>%
  ungroup()
  
```


\newpage

# Results

## Redd estimates

The estimated net error, observed redds and estimates of redds at the reach scale are shown in Table \@ref(tab:redd-est-rch). Plots of the new redds and the GAUC fit to those data are shown in Figure \@ref(fig:gauc-plots). The results are summarized by major areas and population scale in Table \@ref(tab:redd-est-strm). 

```{r redd-est-rch}
redd_results %>%
  mutate(Type = recode(Index,
                       "Y" = 'Index',
                       "N" = "Non-Index")) %>%
  mutate(err_cv = err_se / err_est,
         redd_cv = redd_se / redd_est) %>%
  select(River, Reach, Type, 
         # GAUC,
         `Net Error` = err_est,
         `Net Error CV` = err_cv,
         `Observed Redds ` = tot_feat,
         `Estimated Redds` = redd_est,
         `Std. Err. Redds` = redd_se,
         `Redds CV` = redd_cv) %>%
  kable(digits = c(rep(NA, 3), rep(3, 2), 1, 1, 1, 3),
        format.args = list(drop0trailing = T),
        booktabs = T,
        caption = "Estimates of mean net error and redds for each reach.") %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped", 
                                  "scale_down",
                                  "repeat_header"))
```

```{r gauc-plots, fig.cap = "Plots of observed redd counts (black dots) through time for each qualifying index reach, and the fitted curve from the GAUC model (blue line) with associated uncertainty (gray)."}
redd_results %>%
  filter(GAUC) %>%
  select(River:SurveyType, data) %>%
  unnest(cols = c(data)) %>%
  group_by(River, Reach, Index) %>%
  mutate(survey = 1:n()) %>%
  ungroup() %>%
  ggplot(aes(x = survey, 
             y = NewRedds,
             group = Reach)) +
  geom_point() +
  facet_wrap(~ Reach, 
             scales='free',
             ncol = 2) +
  stat_smooth(method = 'glm', 
              formula = y ~ x + I(x^2), 
              method.args = list(family = quasipoisson), 
              fullrange = T, 
              se = T) + 
  labs(x = 'Survey Number', 
       y = 'New Redds', 
       title='GAUC Reaches') +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  theme(strip.text = element_text(size=12))

```

```{r redd-est-strm}
summ_redd_tab = redd_results %>%
  mutate(Location = if_else(str_detect(Reach,
                                       "^MRW[[:digit:]]",
                                       negate = T),
                            "Tributary",
                            "Lower Methow"),
         Location = factor(Location,
                           levels = c("Lower Methow",
                                      "Tributary"))) %>%
  mutate(across(Reach,
                as.factor)) %>%
  # mutate(Reach = fct_relevel(Reach,
  #                            "W10",
  #                            after = Inf)) %>%
  arrange(River, Location, Reach, Index) %>%
  group_by(River, Location, Index) %>%
  summarize(n_rchs = n(),
            redds_obs = sum(tot_feat),
            redds_est = sum(redd_est),
            redds_se = sqrt(sum(redd_se^2, na.rm = T)),
            redds_cv = redds_se / redds_est,
            .groups = "drop") %>%
  bind_rows(redd_results %>%
              summarise(River = "Total",
                        n_rchs = n(),
                        redds_obs = sum(tot_feat),
                        redds_est = sum(redd_est),
                        redds_se = sqrt(sum(redd_se^2, na.rm = T)),
                        redds_cv = redds_se / redds_est,
                        .groups = "drop"))

summ_redd_tab %>%
  rename(`# Reaches` = n_rchs,
         `Observed Redds ` = redds_obs,
         `Estimated Redds` = redds_est,
         `Std. Err. Redds` = redds_se,
         `Redds CV` = redds_cv) %>%
  kable(digits = c(rep(1, 5), 1, 3),
        booktabs = T,
        format.args = list(drop0trailing = T),
        caption = "Estimate of redds in lower mainstem Methow, and any tributaries surveyed.") %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped",
                                  "scale_down")) %>%
  row_spec(row = nrow(summ_redd_tab) - 1,
           hline_after = T) %>%
  row_spec(row = nrow(summ_redd_tab), 
           bold = T)
  
```

## Spawner Estimates

<!-- Demographic data about sex and origin were derived from the PIT tags detected within each area (Table \@ref(tab:pit-tag-tab)).  -->

```{r pit-tag-tab, eval = F}
met_tags %>% 
  group_by(Location, Sex, Group) %>% 
  summarise(`n tags` = n_distinct(TagID),
            .groups = "drop") %>% 
  kable(booktabs = T,
        caption = "Number of PIT tags detected in each area by sex and origin.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped")
```

Parameter estimates for fish / redd in the lower Methow are shown in Table \@ref(tab:fpr-prop). For the four main areas of the Methow, the number of wild, HOR-SN (hatchery origin safety net) and HOR-C (hatchery origin conservation) tags and their proportions are shown in Table \@ref(tab:hatch-prop). 

```{r fpr-prop}
met_sex_tab %>%
  filter(Location == "Lower Methow") %>%
  select(Area = Location, 
         `F tags` = f_tags,
         `M tags` = m_tags,
         `Fish / redd` = fpr, 
         `FpR Std. Error` = fpr_se) %>%
  kable(digits = 3,
        booktabs = T,
        caption = "Fish per redd estimates.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped")

```

```{r hatch-prop}
met_prop_origin %>%
  kable(digits = 3,
        booktabs = T,
        caption = "Tags and proportion of each origin group, by area.",
        col.names = c('Area', 'Origin Group', 'Total Tags', 'n Tags', 'Proportion', 'Prop Std. Error')) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped")
```

Combining PIT tag-based estimates of spawners in the tributaries with adjusted redd-based estimates of spawners in the lower mainstem Methow, Table \@ref(tab:spawn-est) shows all of them, broken down by area and origin.

```{r spawn-est}
spwn_rch %>%
  mutate(across(Org_Spawners,
                round_half_up)) %>%
  group_by(Location, Origin) %>%
  summarize(Org_Spawners = sum(Org_Spawners),
            Org_SE = sqrt(sum(Org_SE^2)),
            .groups = "drop") %>%
  bind_rows(group_by(., Origin) %>%
              summarize(Org_Spawners = sum(Org_Spawners),
                        Org_SE = sqrt(sum(Org_SE^2)),
            .groups = "drop") %>%
              mutate(Location = "Total")) %>%
  mutate(prnt_val = paste0(round_half_up(Org_Spawners),
                           " (",
                           if_else(Org_SE > 0,
                                   as.character(round(Org_SE / Org_Spawners, 2)),
                                   "-"),
                           ")")) %>%
  select(-starts_with("Org")) %>%
  pivot_wider(names_from = Origin,
              values_from = prnt_val,
              names_sort = T) %>%
  kable(caption = "Estimates (CV) of spawners by area and origin.",
        booktabs = T) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped") %>%
  row_spec(row = nlevels(spwn_rch$Location),
           hline_after = T) %>%
  row_spec(row = nlevels(spwn_rch$Location) + 1, 
           bold = T)
```

## Prespawn Mortality

The estimates of overall prespawn mortality within the Methow population, by origin, are shown in Table \@ref(tab:ps-mort-met). The estimates of prespawn mortality in mainstem areas below Winthrop are displayed in Table \@ref(tab:ps-mort-main). 

```{r ps-mort-met}
ps_mort_met %>%
  mutate(Escapement = paste0(round_half_up(escp), ' (', round_half_up(se), ')'),
         Spawners = paste0(round_half_up(spwn_est), ' (', round_half_up(spwn_se), ')')) %>%
  mutate(ps_mort = if_else(escp < spwn_est, 0, ps_mort),
         ps_mort_se = if_else(escp < spwn_est, NA_real_, ps_mort_se),
         ps_mort_cv = ps_mort_se / ps_mort) %>%
  select(Origin,
         `Initial Escp` = estimate,
         Removals = rem,
         Escapement, Spawners, 
         `Prespawn Mortality` = ps_mort, 
         SE = ps_mort_se,
         CV = ps_mort_cv) %>%
  arrange(desc(Origin)) %>%
  kable(digits = 3,
        booktabs = T,
        caption = "Methow prespawn mortality estimates. Includes estimates (standard error) of initial and final escapement, removals, spawners, prespawn mortality, and standard error of this rate, separated by origin.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped")
```


```{r ps-mort-main}
ps_mort_df %>%
  mutate_at(vars(escp),
            list(~ if_else(. < 0, 0, .))) %>%
  mutate(Escapement = paste0(round_half_up(escp), ' (', round_half_up(se), ')'),
         Spawners = paste0(round_half_up(spwn_est), ' (', round_half_up(spwn_se), ')')) %>%
  mutate(ps_mort = if_else(escp < spwn_est, 0, ps_mort),
         ps_mort_se = if_else(escp < spwn_est, NA_real_, ps_mort_se),
         ps_mort_cv = ps_mort_se / ps_mort) %>%
  select(Origin, Area, 
         `Initial Escp` = estimate,
         Removals = rem,
         Escapement, Spawners, 
         `Prespawn Mortality` = ps_mort, 
         SE = ps_mort_se,
         CV = ps_mort_cv) %>%
  arrange(Area, desc(Origin)) %>%
  kable(digits = 3,
        booktabs = T,
        caption = "Lower mainstem Methow prespawn mortality estimates. Includes estimates (standard error) of escapement, spawners, prespawn mortality, and the standard error of this rate, separated by origin and mainstem areas below Winthrop.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped")
```



# Discussion

Most of the covariates collected in `r yr` were within the range of those in the model data set from @Murdoch2018, leading to estimates of net error in a very similar range to the model dataset (Figure \@ref(fig:covariate-comparison)). However, some reaches did not meet the minimum thresholds of number of observed redds or number of weeks with at least one new redd observed, so we used the GAUC method in only `r as.character(as.english(sum(redd_results$GAUC)))` reaches. 

<!-- Estimates of prespawn mortality were lower for the Methow overall compared to only the mainstem areas (TableS \@ref(tab:ps-mort-met) and \@ref(tab:ps-mort-main)). This reflects our assumption that steelhead may overwinter in the lower mainstem Methow for a considerable period of time, but when they move into the tributaries they are ready to spawn and therefore PIT tag estimates of escapement to those tributaries are also estimates of spawners in those tributaries.  -->


<!-- The estimates of high prespawn mortality in the lower mainstem of the Methow for hatchery fish could be accurate, but it should be noted that many of the redd surveys failed to observe a single redd in many of the reaches (Table \@ref(tab:redd-est-rch)). Without any observed redds, any estimate of net error is moot, as the adjusted redd estimate will still be zero. So if all the redds were missed in some of those reaches, the estimate of total spawners in the lower mainstem should be higher, leading to a lower estimate of prespawn mortality. It is unclear whether that actually occurred, or if there were actually no redds this year in those reaches. -->

Some of the estimates of prespawn mortality were zero this year, due to higher estimates of spawners compared to escapement, at least after removals had been accounted for. In some cases, there were overlapping confidence intervals between spawners and escapement, so not too much should be made of that fact, and we interpret that as at least very low levels of prespawn mortality, perhaps even none. It should be noted that our assumption of no pre-spawn mortality once fish enter the tributaries (i.e. after moving out of the lower mainstem Methow) may be violated. For example, it appears there is at least several weeks between the earliest PIT tag detection and peak spawning. With the existing data, it can be difficult to estimate pre-spawn mortality in these tributaries.

In addition, in 2021, because of the very skewed sex ratio, many of the male steelhead caught in the lower mainstem were live spawned and then returned to the river. This could mean that the escapement estimates in Tables \@ref(tab:ps-mort-met) and \@ref(tab:ps-mort-main) should perhaps be a bit higher, even as high as the estimated number of spawners. 


# Acknowledgements

The data for this report was collected by Washington Department of Fish and Wildlife. Development of the observer error model was done in collaboration with Andrew Murdoch, WDFW.

<!-- The following line inserts a page break  -->
\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

<!-- \newpage -->

<!-- ### Colophon -->

<!-- This report was generated on `r Sys.time()` using the following computational environment and dependencies:  -->

<!-- ```{r colophon, cache = FALSE} -->
<!-- # which R packages and versions? -->
<!-- if ("devtools" %in% installed.packages()) devtools::session_info() -->
<!-- ``` -->

<!-- The current Git commit details are: -->

<!-- ```{r} -->
<!-- # what commit is this file at?  -->
<!-- if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())   -->
<!-- ``` -->
