---
title: "Estimates of Wenatchee Steelhead Redds and Spawners"
subtitle: "Spawn Year 2020"
author:
  - Kevin See:
      email: Kevin.See@merck.com
      institute: [biomark]
      correspondence: true
institute:
  - biomark: Biomark, Inc.
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      fig_height: 6
      fig_width: 6
      toc: yes
      toc_depth: 3
      toc_float:
        collapsed: yes
        smooth_scroll: yes
      theme: simplex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::pdf_document2:
      fig_caption: yes
      fig_height: 5
      fig_width: 6
      toc: yes
      includes:
        in_header: ../templates/header_ABS.tex
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      fig_height: 4
      fig_width: 6
      toc: yes
      reference_docx: "../templates/ReportTemplate.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography:
- packages.bib
- references.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  This report contains estimates of total steelhead spawners in the Wenatchee.
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights.
---


<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

```{r packages}

# load these packages
library(tidyverse)
library(lubridate)
library(janitor)
library(magrittr)
library(msm)
library(knitr)
library(kableExtra)
library(english)
library(UCSthdReddObsErr)
library(here)

theme_set(theme_bw())

options(knitr.kable.NA = '-')

# when knitting to Word, use this
# what kind of document is being created?
doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

if(doc.type == 'docx') {
  options(knitr.table.format = "pandoc")
}

```

```{r package-bibtex, eval = F}
knitr::write_bib(c("tidyverse", 
                   "lubridate", 
                   "janitor", 
                   "knitr",
                   "magrittr",
                   "msm"),
                 file = 'packages.bib')
```


```{r load-data}
# what year is this report for?
yr = 2020

# load data for this year
load(here('analysis/data/derived_data',
          paste0('wen_', yr, '.rda')))

# org_data = read_rds('../data/derived_data/two_obs_original_data.rds')
# net_err_mod = read_rds('../data/derived_data/two_obs_net_error_model.rds')


# how many fish removed this year?
data("removal_df")
rem_df = removal_df %>%
  filter(Year == yr) %>%
  select(-Year)
```

```{r rt-survival}
# data from WDFW (Nate Fuchs' radio telemetry study)
rt_df = tibble(year = rep(2015:2016, each = 2),
               Origin = rep(c("Hatchery", "Natural"), 2),
               ow_fish = c(20, 25, 4, 12),
               surv_fish = c(16, 24, 3, 9)) %>%
  mutate(phi = surv_fish / ow_fish,
         phi_se = sqrt((phi * (1 - phi))/ ow_fish))

# add years together
rt_df %<>%
  bind_rows(rt_df %>%
              group_by(Origin) %>%
              summarize(across(c(ow_fish, surv_fish),
                               sum),
                        .groups = "drop") %>%
              mutate(phi = surv_fish / ow_fish,
                     phi_se = sqrt((phi * (1 - phi))/ ow_fish))) %>%
  mutate(across(year,
                as.factor)) %>%
  mutate(across(year,
                fct_explicit_na,
                na_level = "Total"))
```

```{r}
spwn_df <- escp_wen %>%
  ungroup() %>%
  filter(Area == "Wen_all") %>%
  left_join(rem_df %>%
              group_by(Origin) %>%
              summarize(across(rem,
                               sum))) %>%
  rowwise() %>%
  mutate(escp = max(0, estimate - rem)) %>%
  ungroup() %>%
  full_join(rt_df %>%
              group_by(Origin) %>%
              summarize(across(c(ow_fish, surv_fish),
                               sum),
                        .groups = "drop") %>%
              mutate(phi = surv_fish / ow_fish,
                     phi_se = sqrt((phi * (1 - phi))/ ow_fish)) %>%
              select(-ends_with("fish"))) %>%
  rowwise() %>%
  mutate(all_spwn = escp * phi) %>%
  mutate(all_spwn_se = msm::deltamethod(~ x1 * x2,
                                        mean = c(escp, phi),
                                        cov = diag(c(se, phi_se)^2))) %>%
  ungroup() %>%
  left_join(trib_spawners %>%
              group_by(Origin) %>%
              summarize(across(Spawners,
                               sum),
                        across(Spawners_SE,
                               ~ sqrt(sum(.^2))),
                        .groups = "drop") %>%
              rename(trib_spwn = Spawners,
                     trib_se = Spawners_SE)) %>%
  mutate(main_spwn = all_spwn - trib_spwn,
         main_spwn_se = sqrt(all_spwn_se^2 + trib_se^2),
         across(main_spwn,
                ~ if_else(. < 0, 0, .)))

spwn_main = spwn_df %>%
  mutate(Location = "Mainstem") %>%
  select(Origin,
         Location,
         Spawners = main_spwn,
         Spawners_SE = main_spwn_se)
```


```{r spawner-estimate}
# spwn_main = escp_wen %>%
#   filter(Area != "Wen_all") %>%
#   ungroup() %>%
#   left_join(rem_df %>%
#               select(-Source)) %>%
#   rowwise() %>%
#   mutate(escp = max(0, estimate - rem)) %>%
#   ungroup() %>%
#   full_join(rt_df %>%
#               group_by(Origin) %>%
#               summarize(across(c(ow_fish, surv_fish),
#                                sum),
#                         .groups = "drop") %>%
#               mutate(phi = surv_fish / ow_fish,
#                      phi_se = sqrt((phi * (1 - phi))/ ow_fish)) %>%
#               select(-ends_with("fish"))) %>%
#   rowwise() %>%
#   mutate(Spawners = escp * phi) %>%
#   mutate(Spawners_SE = msm::deltamethod(~ x1 * x2,
#                                         mean = c(escp, phi),
#                                         cov = diag(c(se, phi_se)^2))) %>%
#   ungroup() %>%
#   rename(Location = Area)

spwn_strm = spwn_main %>%
  select(any_of(names(trib_spawners))) %>%
  bind_rows(trib_spawners) %>%
  mutate(Location = recode(Location,
                           'past_CHL' = 'Chiwawa',
                           'past_CHM' = 'Chumstick',
                           'past_CHW' = 'Chiwaukum',
                           'past_ICL' = 'Icicle',
                           'past_LWN' = 'Little Wenatchee',
                           'past_MCL' = 'Mission',
                           'past_NAL' = 'Nason',
                           'past_PES' = 'Peshastin',
                           'past_WTL' = 'White River'),
         Location = factor(Location, levels = c("Mainstem",
                                                # 'Below_TUM',
                                                # 'TUM_bb',
                                                'Icicle', 
                                                'Peshastin', 
                                                'Mission', 
                                                'Chumstick', 
                                                'Chiwaukum', 
                                                'Chiwawa', 
                                                'Nason', 
                                                'Little Wenatchee', 
                                                'White River'))) %>%
  arrange(Location, Origin)

# spwn_strm %>%
#   pivot_wider(id_cols = "Location",
#               names_from = "Origin",
#               values_from = c('Spawners',
#                               'Spawners_SE'))

```


\newpage

# Introduction

Redd counts are an established method to provide an index of adult spawners [@Gallagher2007]. However, all redd surveys were cancelled during the steelhead spawning season due to COVID-19. Therefore, for this year, we have devised another method to estimate the number of spawners In the Wenatchee subbasin, based on results from a PIT-tag based patch-occupancy model that estimates escapement, and a radio telemetry study that estimated overwinter survival in the mainstem of the Wenatchee.

# Methods

## Data

### Escapement

Estimates of escapement to various tributaries in the Wenatchee were made using a branching patch-occupancy model [@Waterhouse2020] based on PIT tag observations of fish tagged at Priest Rapids dam. All fish that escaped to the various tributaries were assumed to be spawners (i.e. prespawn mortality only occurs in the mainstem). The remaining fish may have survived the winter and spawned in the mainstem (where redd counts would normally be conducted). 

### Overwinter survival

Estimates of overwinter survival in the Wenatchee subbasin come from a radio telemetry study, conducted over two years (2015 and 2016) in the Wenatchee [@Fuchs_review]. Steelhead in the study were both radio and PIT tagged, and zero mortality was observed in fish once they entered the tributaries of the Wenatchee. Tags were combined across both years of the study, since we are making the assumption that overwinter survival is consistent year to year. 

```{r rt-table}
rt_df %>%
  rename(Year = year,
         `Initial Fish` = ow_fish,
         `Surviving Fish` = surv_fish,
         Survival = phi,
         SE = phi_se) %>%
  kable(booktabs = T,
        linesep = "",
        digits = 2,
        caption = "Data from the radio telemetry study, representing fish known to have escaped to the Wenatchee River.") %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped")) %>%
  row_spec(row = (nrow(rt_df)-1):nrow(rt_df),
           bold = T) %>%
  row_spec(row = nrow(rt_df) - 2,
           hline_after = T)
```

### Known removals

Before applying overwinter survivals, we must account for any fish removed at Tumwater or Dryden for brookstock or surplus, as well as any deaths due to harvest (Table \@ref(tab:removals)). 

```{r removals}
rem_df %>%
  rename(Removed = rem) %>%
  select(-Area) %>%
  spread(Origin, Removed) %>%
  kable(caption = "Known number of fish removed at dams or due to harvest, by origin.",
        booktabs = T) %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped"))
```

## Analysis

Due to a lack of redd counts in 2020, we instead focus on estimates of escapement from the PIT-tag based patch-occupancy model, by origin. We started with escapement to the entire Wenatchee subbasin (past site LWE), and then subtracted known removals. After this, we applied an overwinter survival estimate to determine how many fish survived to spawn. We determined how many mainstem spawners by subtracting the total number of estimated tributary spawners from the total population level estimate of spawners. 

\newpage

# Results

## Mainstem spawners

Estimates of mainstem spawners above and below Tumwater dam are shown in Table \@ref(tab:mainstem-spwn). 

```{r mainstem-spwn}
spwn_df %>%
  mutate(Location = "Mainstem") %>%
  # mutate(Location = recode(Location,
  #                          "Below_TUM" = "Below TUM",
  #                          "TUM_bb" = "Above TUM")) %>%
  mutate(across(main_spwn,
                round_half_up)) %>%
  select(Location,
         Origin,
         `PO Estimate` = estimate,
         `PO SE` = se,
         `Removed` = rem,
         Escapement = escp,
         `Overwinter Surv.` = phi,
         `Overwinter SE` = phi_se,
         `Trib Spawners` = trib_spwn,
         `Trib SE` = trib_se,
         Spawners = main_spwn,
         `Spawners SE` = main_spwn_se) %>%
  kable(caption = "Inputs and estimates of mainstem spawners.",
        linesep = "",
        digits = c(0, 0, 
                   0, 1, 
                   0, 0, 
                   3, 3, 
                   0, 1,
                   0, 1),
        booktabs = T) %>%
  kable_styling(bootstrap_options = c("striped", 
                                      "hover"),
                latex_options = c("striped",
                                  "scale_down"))
```

## Total spawners

Table \@ref(tab:spawn-est) displays estimates of spawners in all areas within the Wenatchee, as well as the total by origin.

```{r spawn-est}
spwn_strm %>%
  # mutate(Location = recode(Location,
  #                          "Below_TUM" = "Below TUM",
  #                          "TUM_bb" = "Above TUM")) %>%
  # bind_rows(spwn_strm %>%
  #             group_by(Origin) %>%
  #             summarize(Spawners = sum(round_half_up(Spawners)),
  #                       Spawners_SE = sqrt(sum(Spawners_SE^2)),
  #                       .groups = "drop") %>%
  #             mutate(Location = "Total")) %>%
  bind_rows(spwn_df %>%
              mutate(Location = "Total") %>%
              mutate(across(all_spwn,
                            round_half_up)) %>%
              select(Origin,
                     Location,
                     Spawners = all_spwn,
                     Spawners_SE = all_spwn_se)) %>%
  mutate(spwn_cv = Spawners_SE / Spawners,
         spwn_cv = if_else(spwn_cv == Inf, NA_real_, spwn_cv)) %>%
  mutate(prnt_val = paste0(round_half_up(Spawners), " (", if_else(is.na(spwn_cv), '-', as.character(round(spwn_cv, 2))), ")")) %>%
  select(Location, Origin, prnt_val) %>%
  pivot_wider(names_from = "Origin",
              values_from = "prnt_val") %>%
  select(Area = Location,
         Natural,
         Hatchery) %>%
  kable(caption = "Estimates (CV) of spawners by area and origin.",
        booktabs = T) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                latex_options = "striped") %>%
  row_spec(row = nrow(spwn_strm)/2,
           hline_after = T) %>%
  row_spec(row = nrow(spwn_strm)/2 + 1, 
           bold = T)
```


# Discussion

Despite the lack of redd count data this year, we were able to estimate the number of spawners, by origin, through applying an empirical estimate of overwinter survival to total escapement estimates to the subbasin. After accounting for known removals and tributary spawners, the estimates of mainstem spawners for hatchery origin fish did drop below zero. Therefore, we fixed the estimate of mainstem hatchery spawners to zero. 

# Acknowledgements

The data for this report was collected by Washington Department of Fish and Wildlife. Development of the observer error model was done in collaboration with Andrew Murdoch, WDFW.

<!-- The following line inserts a page break  -->
\newpage

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

<!-- \newpage -->

<!-- ### Colophon -->

<!-- This report was generated on `r Sys.time()` using the following computational environment and dependencies:  -->

<!-- ```{r colophon, cache = FALSE} -->
<!-- # which R packages and versions? -->
<!-- if ("devtools" %in% installed.packages()) devtools::session_info() -->
<!-- ``` -->

<!-- The current Git commit details are: -->

<!-- ```{r} -->
<!-- # what commit is this file at?  -->
<!-- if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())   -->
<!-- ``` -->
